---
title: "Bank Marketing Case Study 1"
author: "Dominica Peri"
date: "2025-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Executive Summary (Last)

2. The Problem
    1. Introduction/Background
    2. Purpose of study/importance of study/statement of problem
    3. Questions to be answered/conceptual statement of hypotheses
    4. Outline of remainder of report (brief)
    
3. Review of Related Literature
    1. Acquaint reader with existing methodologies used in this area
    
4. Methodology
    1. Identification, classification and operationalization of variables.
    2. Statements of hypotheses being tested and/or models being developed.
    3. Sampling techniques, if full data is not being used.
    4. Data collection process, including data sources, data size, etc. Primary/secondary?
    5. Modeling analysis/techniques used
    6. Methodological assumptions and limitations.
    
5. Data
    1. Data cleaning
    2. Data preprocessing
    3. Data limitations
    
6. Findings (Results)
    1. Results presented in tables or charts when appropriate
    2. Results reported with respect to hypotheses/models.
    3. Factual information kept separate from interpretation, inference and evaluation.
    
7. Conclusions and Recommendations
    1. Discuss alternative methodologies

#____________________________________________________________________________

Load Libraries

```{r Libraries}
library(MASS) 
library(ggplot2)
library(tidyverse)
library(corrplot)
library(car)
library(GGally)
library(ROCR)
library(data.table)
library(Hmisc)
library(tibble)
library(knitr)
library(here)
```

Import Data

```{r Bank Data}
data_path <- "https://raw.githubusercontent.com/kccasey1985/Case-Study-Bank-Marketting-/main/bank-additional-full.csv"

df <- fread(data_path, sep = ";") |> as_tibble()
glimpse(df)
```
```{r Bank Data}
opts_knit$set(root.dir = here())

# load the dataset
df <- fread(here("bank-additional-full.csv"), sep = ";") |> as_tibble()
glimpse(df)
```


```{r Bank Data}

head(df)

```


```{r Bank Data}

str(df)

```

```{r Bank Data}

summary(df)

```

```{r Bank Data}

df2 <- df |> 
  mutate(job = factor(job),
         marital  = factor(marital),
         education  = factor(education),
         default  = factor(default),
         loan  = factor(loan),
         contact  = factor(contact),
         month   = factor(month),
         day_of_week  = factor(day_of_week),
         poutcome    = factor(poutcome),
         y   = factor(y),
         housing  = factor(housing))



```

```{r Bank Data}

str(df2)

```

```{r Bank Data}

summary(df2)

```
```{r Bank Data}

anyNA(df2)

```

```{r Bank Data}

df2 <- df2 |> 
  mutate(age= cut(age, breaks = c(20, 30, 40, 50, 60, 70, 100),
                         labels = c("21-30", "31-40", "41-50", "51-60", "61-70", "71+"),
                         right = F))|>
  select(-duration)

```

```{r Bank Data}

summary(df2)

```


```{r Bank Data}

df2_num <- dplyr::select_if(df2, is.numeric)
matrix = cor(df2_num)
corrplot(matrix, method = c("number")) 

```







```{r Bank Data}

corrplot::corrplot(matrix, method = "circle")


```





```{r Bank Data}

lm_model <- lm(y ~ ., data = df2)

```
```{r}

# Make sure y is a factor with two levels
df2$y <- factor(df2$y, levels = c("no", "yes"))

# Logistic regression
logit_model <- glm(y ~ ., data = df2, family = binomial)

summary(logit_model)

```
```{r}
alias(lm_model)

```

`



```{r Bank Data}
vif_values <- car::vif(lm_model)
print(vif_values)


```


